---
title: "Group 23"
date: "9/11/2022"
output: html_document
---

```{r}
library(plyr)
library(dplyr)
library(stringr)
library(glmnet)
library(ggplot2)
library(lubridate)
```

# Question 1 & 2
```{r}
getcvplot <- function(N_industry, date){
    data <- na.omit(read.csv('NEW_48_Industry_Portfolios_Daily.csv', na.strings = c("-99.99"), skip = 9)) #25292
    Nindustry <- length(N_industry)
    columns <- colnames(data)[c(1, N_industry)] # Take out the columns of the chosen industry
    w_ew = matrix(1/Nindustry, nrow = Nindustry, ncol = 1) # Get the equally weighted matrix
    N = rbind(diag(Nindustry - 1), c(rep(-1, Nindustry - 1))) # N as identified in the mathematical equation
    # Number of rows to collect = 63 to estimate the covariance matrix E
    end_row <- which(data$X == date) # To identify the row that correspond to the date
    start_row <- end_row - 63 + 1 # To get the starting data to collect for the 63 observations
    data1 <- data[columns][start_row:end_row,] # 63 observations
    demean_data <- scale(data1[-1], scale = F) # Scale the data such that the data is demean where the sample mean of each asset return is 0
    y = demean_data %*% w_ew # Matrix multiplication of Rwew
    X = demean_data %*% N # Matrix multiplication of RN
    lasso = cv.glmnet(X,y, alpha = 1, standardise = F, intercept = F) #1 represents lasso, no requirement to have standardization
    ridge = cv.glmnet(X,y, alpha = 0, standardise = F, intercept = F) #0 represents ridge, no requirement to have standardization
    layout(matrix(c(1,2), 2, 2, byrow = TRUE))
    plot(lasso, sign.lambda = -1, main = paste(Nindustry, "Industries in Lasso"))
    plot(ridge, sign.lambda = -1, main = paste(Nindustry, "Industries in Ridge"))
}
```

```{r}
# 6 Industries
# N_industry = first 6 industries 
Industry_6 <- c(2:7)
getcvplot(Industry_6, 20220103)

# 24 industries
# N_industry = first 24 industries
Industry_24 <- c(2:25)
getcvplot(Industry_24, 20220103)

# 48 industries
Industry_48 <- c(2:49)
getcvplot(Industry_48, 20220103)
```

# Question 4

#loading the File
```{r}
df <- read.csv('NEW_48_Industry_Portfolios_Daily.csv', skip = 9, na.strings = "-99.99")

colnames(df)[1] <- 'Dates' #rename dates column

df$Dates <- as.character(df$Dates)
df$row_number <- seq.int(nrow(df)) #generate row numbers

dates_row_number <- df[which(df$Dates == '20200601'):which(df$Dates == '20220630'),] %>% select(Dates, row_number)

df <- df %>% select(-row_number) #remove row_number column

######################################################################
############### Input number of industries here ######################

num_of_industry <- 48

######################################################################
######################################################################

################### NOTE!!!!! ########################################
#after which, we can run all the following chunks of code to generate a cumulative return plot for lasso, ridge, minvar and ew model as a png file
#to generate different number of industries, amend variable 'num_of_industry' above to desired number, then re-run all the following chunks of code below
```

#Cummulative Return Code
##function to derive weights of industries
```{r}
get_weights <- function(initial_df, number_of_industry, row_number){
  # initial_df = df
  # number_of_industry = 48
  # row_number = 24897
  ##this function returns a list of weights for lasso, ridge and minvar for 63 observations prior to a certain date
  
  #specify number of industry
  N_industry <- c(2:(number_of_industry+1))
  #specify columns
  columns <- colnames(initial_df)[c(1, N_industry)]
  w_ew <- matrix(1/length(N_industry), nrow = length(N_industry), ncol = 1) # Get the equally weighted matrix
  
  # N identity matrix as identified in the mathematical equation
  N <- rbind(diag(length(N_industry) - 1), matrix(-1, ncol = length(N_industry) - 1)) 
  
  # To get the starting data to collect for the 63 observations
  end_row <- row_number - 1 
  start_row <- end_row - 63 + 1
  
  #slice data to obtain 48 industries with respective 63 daily returns before that date
  data1 <- initial_df[columns][start_row:end_row,]
  
  # Scale the data such that the data is demean where the sample mean of each asset return is 0
  # by specifying center=T and scale=F, we aim to subtract the mean from each individual return but NOT dividing by the standard deviation.
  demean <- scale(data1[, c(2:length(columns))], center=T, scale = F) 
  
  y <- demean %*% w_ew # Matrix multiplication of RWew
  x <- demean %*% N # Matrix multiplication of RN
  lasso <- cv.glmnet(x, y, alpha = 1, standardise = F, intercept = F) 
  ridge <- cv.glmnet(x, y, alpha = 0, standardise = F, intercept = F) 
  minvar <- lm(y ~ x)
  
  #applying weights of coefficients.
  lasso_beta <- t(matrix(coef(lasso, s = "lambda.min")))[-1] #lambda.min corresponds to the lowest MSE
  ridge_beta <- t(matrix(coef(ridge, s = "lambda.min")))[-1]
  minvar_beta <- t(matrix(coef(minvar)))[-1]
  
  lasso_weights <- w_ew - N %*% lasso_beta
  ridge_weights <- w_ew - N %*% ridge_beta
  minvar_weights <- w_ew - N %*% minvar_beta
  
  #returns a list of lasso_weights, ridge_weights and minvar_weights
  return (
    list(
      lasso_weights,
      ridge_weights,
      minvar_weights,
      w_ew
    )
    )
}
```

#get weights functions
```{r}
#the following functions return the weights of lasso, ridge and minvar respectively from 'get_weights' function
get_lasso_weights <- function(day_weights_lst){
  return (list(as.vector(day_weights_lst[[1]])))
}

get_ridge_weights <- function(day_weights_lst){
  return (list(as.vector(day_weights_lst[[2]])))
}

get_minvar_weights <- function(day_weights_lst){
  return (list(as.vector(day_weights_lst[[3]])))
}

get_ew_weights <- function(day_weights_lst){
  return (as.vector(day_weights_lst[[4]]))
}

```

#list of weights for each model
```{r}
#initialise vector of betas for each model
lasso_weights_lst <- c() 
ridge_weights_lst <- c()
minvar_weights_lst <- c()

#for loop to derive a list of set of weights of industries for every date within data
start_date_index <- dates_row_number$row_number[1]
end_date_index <- dates_row_number$row_number[nrow(dates_row_number)]

start_time <- Sys.time()

for (i in start_date_index:end_date_index){
  day_weights_lst <- get_weights(df, num_of_industry, i)
  lasso_weights_lst <- append(lasso_weights_lst, get_lasso_weights(day_weights_lst)) #append lasso weights for the date
  ridge_weights_lst <- append(ridge_weights_lst, get_ridge_weights(day_weights_lst))#append ridge weights for the date
  minvar_weights_lst <- append(minvar_weights_lst, get_minvar_weights(day_weights_lst)) #append minvar weights for the date
}

end_time <- Sys.time()
elapsed_time <- end_time - start_time
elapsed_time
```

#out of sample day return function
```{r}
#this function returns the out of sample return of a day given the weights of the portfolio derived from the past 63 observations
daily_returns <- function(weights_lst, initial_df, dates_df, first_date_index, num_of_industry){
  #initialise date index
  date_index = first_date_index
  #function to return vector of daily returns
  #initialise vector of returns
  returns <- c()
  #for loop to generate returns from lasso regularised model
  for (i in 1:nrow(dates_df)){
    #sumproduct weights and individual returns
    day_return <- crossprod(weights_lst[[i]], as.numeric(as.vector(df[date_index, 2:(num_of_industry + 1)])))
    #append weighted return
    returns <- append(returns, day_return)
    date_index <- date_index + 1 #increase date index
  }
  return (returns)
}

daily_ew_returns <- function(equally_weighted_matrix, initial_df,  dates_df, first_date_index, num_of_industry){
  #initialise date index
  date_index = first_date_index
  #function to return vector of daily returns
  #initialise vector of returns
  returns <- c()
  #for loop to generate returns from lasso regularised model
  for (i in 1:nrow(dates_df)){
    #sumproduct weights and individual returns
    day_return <- crossprod(equally_weighted_matrix, as.numeric(as.vector(df[date_index, 2:(num_of_industry + 1)])))
    #append weighted return
    returns <- append(returns, day_return)
    date_index <- date_index + 1 #increase date index
  }
  return (returns)
}
```

#create vector of daily returns and initialise vector of cumulative returns
```{r}

#equally weighted matrix
w_ew <- get_ew_weights(get_weights(df, num_of_industry, start_date_index))

lasso_daily_returns <- daily_returns(lasso_weights_lst, df, dates_row_number, start_date_index, num_of_industry)
ridge_daily_returns <- daily_returns(ridge_weights_lst, df, dates_row_number, start_date_index, num_of_industry)
ew_daily_returns <- daily_ew_returns(w_ew, df, dates_row_number, start_date_index, num_of_industry)
minvar_daily_returns <- daily_returns(minvar_weights_lst, df, dates_row_number, start_date_index, num_of_industry)

#write.csv(ew_daily_returns, "ew_daily_returns.csv")

#initialise vector of cumulative returns with initial investment of $1 for both ridge and lasso
cum_lasso_returns_vector <- c(1* (1 + lasso_daily_returns[1] / 100))
cum_ridge_returns_vector <- c(1 * (1 + ridge_daily_returns[1] / 100))
cum_ew_returns_vector <- c(1 * (1 + ew_daily_returns[1] / 100))
cum_minvar_returns_vector <- c(1 * (1 + minvar_daily_returns[1] / 100))
```

#cumulative returns function
```{r}
cum_return <- function(dates_df, return_vector, cum_return_vector){
  #function returns cum return for initial investment of $1
  #inputs 'return vector' is the vector of daily returns
  for (i in 2:nrow(dates_df)){
    # return of that day
    return <- 1 + return_vector[i] / 100
    # new cum return based on previous day daily return
    new_cum_r <- cum_return_vector[i - 1] * return
    # append new cum return
    cum_return_vector <- append(cum_return_vector, new_cum_r)
  }
  return (cum_return_vector)
}
```

## creating df of returns for all models
```{r}
lasso_cum_return <- cum_return(dates_row_number, lasso_daily_returns, cum_lasso_returns_vector) #lasso regularised cum return
ridge_cum_return <- cum_return(dates_row_number, ridge_daily_returns, cum_ridge_returns_vector) #ridge regularised cum return
ew_cum_return <- cum_return(dates_row_number, ew_daily_returns, cum_ew_returns_vector) #equally weighted cum return
minvar_cum_return <- cum_return(dates_row_number, minvar_daily_returns, cum_minvar_returns_vector) #equally weighted cum return

#create returns dataframe
returns_df <- dates_row_number %>% 
  select(Dates) %>%
  mutate(Dates = ymd(Dates),
         Ridge_Cum_Return = ridge_cum_return,
         Lasso_Cum_Return = lasso_cum_return,
         EW_Cum_Return = ew_cum_return,
         MinVar_Cum_Return = minvar_cum_return)
```

##plotting graph and saving as .png file
```{r}
plt <- ggplot(returns_df, aes(x = Dates)) +
  geom_line(aes(y = Ridge_Cum_Return, color = "Ridge")) +
  geom_line(aes(y = Lasso_Cum_Return, color = "LASSO")) +
  geom_line(aes(y = EW_Cum_Return, color = "Equally-Weighted")) +
  geom_line(aes(y = MinVar_Cum_Return, color = "MinVar")) +
  scale_color_manual(name = "Legend", 
                     values = c(
                       "Ridge" = "#146FC2", 
                       "LASSO" = "#4fa861", 
                       "Equally-Weighted" = "#2A363B",
                       "MinVar" = "#db3c30"
                       )
                     ) +
  labs(title = paste("Cumulative Return for ", num_of_industry, " industries"), 
       subtitle = "based on $1 initial investment", 
       y = 'Portfolio Value', 
       x = "Dates") +
  theme(plot.title = element_text(face="bold"),
        plot.subtitle = element_text(face = "italic")) +
  theme_bw()

ggsave(filename = paste(num_of_industry, " industries cum return.png"), 
       plot = plt, 
       height = 7,
       width = 12)

```
